Extended Data

### 2022-02-21 doc/manuscript/Dow_manuscript_Nature.Rmd
new line for each sentence
This makes it easier to track changes

Our finding that interannual variation in stem growth is more strongly linked to conditions during the peak growing season than to growing season length aligns with recent studies on annual stem growth[@etzold_number_2022; @tumajer_limitation_2022;@delpierre_wood_2016] and with parallel findings for _NEE_[@zhou_explaining_2016; @fu_maximum_2019].
However, there is also a disconnect with findings that _NEE_ is at least modestly greater in years with warm springs[@keenan_net_2014] or long growing seasons[@zhou_explaining_2016; @fu_maximum_2019;@finzi_carbon_2020].
Warming advances spring phenology and may advance or delay autumn senescence depending on timing of warming and water availability[@zohner_ongoing_2019;@xie_predicting_2018; @crabbe_extreme_2016], with delays more common across eastern North America [@buermann_widespread_2018;@finzi_carbon_2020;@keenan_net_2014], implying that warming temperatures are lengthening the period from peak stem growth to the cessation of CO~2~ uptake by the ecosystem.
Cite delpierre 2016 around here. Delpierre et al. 2016 found that stem growth was not correlated with carbon inputs, agreeing with the results here. However, the authors 
We show that the extra C fixation in years with warm springs does not substantially augment stem growth, but it remains unclear how it is allocated within the ecosystem.
There are two main possibilities, which hold contrasting implications for the response of forest C balance to rising spring temperatures.

@@

Canopy foliage phenology data for the years 2001-2018 were extracted for SCBI and Harvard Forest from the MCD12Q2 V6 Land Cover Dynamics product (a.k.a. MODIS Global Vegetation Phenology product)[@friedl_mcd12q2_2019] via Google Earth Engine. For each year at each site, we extracted data from the pixel containing the center of each forest plot (resolution of 500m). Using the daily MODIS 2-band Enhanced Vegetation Index data (EVI2) , the product yields the timing of phenometrics (vegetation phenology) over each year, including timing of greenup, midgreenup, and senescence as used in this study. Data points were kept if they were flagged as "good" or "best" quality.

>>>>

Our finding that interannual variation in stem growth is more strongly linked to conditions during the peak growing season than to growing season length aligns with recent studies on annual stem growth[@etzold_number_2022; @tumajer_limitation_2022;@delpierre_wood_2016] and with parallel findings for _NEE_[@zhou_explaining_2016; @fu_maximum_2019].
However, there is also a disconnect with findings that _NEE_ is at least modestly greater in years with warm springs[@keenan_net_2014] or long growing seasons[@zhou_explaining_2016; @fu_maximum_2019;@finzi_carbon_2020].
Warming advances spring phenology and may advance or delay autumn senescence depending on timing of warming and water availability[@zohner_ongoing_2019;@xie_predicting_2018; @crabbe_extreme_2016], with delays more common across eastern North America [@buermann_widespread_2018;@finzi_carbon_2020;@keenan_net_2014], implying that warming temperatures are lengthening the period from peak stem growth to the cessation of CO~2~ uptake by the ecosystem.
**(Cite delpierre 2016 around here.) Delpierre et al. 2016 found that stem growth was not correlated with carbon inputs, agreeing with the results here. However, the authors ...**
We show that the extra C fixation in years with warm springs does not substantially augment stem growth, but it remains unclear how it is allocated within the ecosystem.
There are two main possibilities, which hold contrasting implications for the response of forest C balance to rising spring temperatures.

@@

Canopy foliage phenology data for the years 2001-2018 were extracted for SCBI and Harvard Forest from the MCD12Q2 V6 Land Cover Dynamics product (a.k.a. MODIS Global Vegetation Phenology product)[@friedl_mcd12q2_2019] via Google Earth Engine. 
For each year at each site, we extracted data from the pixel containing the center of each forest plot (resolution of 500m). 
Using the daily MODIS 2-band Enhanced Vegetation Index data (EVI2), the product yields the timing of phenometrics (vegetation phenology) over each year, including timing of greenup, midgreenup, and senescence as used in this study. 
Data points were kept if they were flagged as "good" or "best" quality.
